Comparing gbm vs rf (default parameterisations)

### 2020-11-09 src/s_cauchy.Rmd
cleaning up repo and rerunning rhat experiments

mon <- monitor_extra(x)
rhat <- mon$zfsRhat
rhat1 <- mon$zsRhat
rhat <- tibble(rhat, rhat1) %>% 
  apply(1, max)

@@

mon <- monitor_extra(full_data)
rhat <- mon$zfsRhat
rhat1 <- mon$zsRhat
rhat <- tibble(rhat, rhat1) %>% 
  apply(1, max)
a_rhat_df <-a_rhat_df %>% 
  mutate(alt=rhat[101:150],
         source="r_hat")
```

@@

r_star_nom_gbm <- r_star(full_data, method="gbm", uncertainty = T)
r_star_nom_rf <- r_star(full_data, method="rf", uncertainty = T)
mon_nom <- monitor_extra(full_data)

rhat <- mon_nom$zfsRhat
rhat1 <- mon_nom$zsRhat
rhat_nom <- tibble(rhat, rhat1) %>% 
  apply(1, max)
```

@@

r_star_alt_gbm <- r_star(full_data, method="gbm", uncertainty = T)
r_star_alt_rf <- r_star(full_data, method="rf", uncertainty = T)
mon_alt <- monitor_extra(full_data)
rhat <- mon_alt$zfsRhat
rhat1 <- mon_alt$zsRhat
rhat_alt <- tibble(rhat, rhat1) %>% 
  apply(1, max)

>>>>

mon <- monitor_extra(x)

# calculate rhat as in posterior package
rhat <- mon$zfsRhat
rhat1 <- mon$zsRhat
rhat <- tibble(rhat, rhat1) %>% 
  apply(1, max)

@@

# calculate rhat as in posterior package
mon <- monitor_extra(full_data)
rhat <- mon$zfsRhat
rhat1 <- mon$zsRhat
rhat <- tibble(rhat, rhat1) %>% 
  apply(1, max)
a_rhat_df <-a_rhat_df %>% 
  mutate(alt=rhat[101:150],
         source="r_hat")
```

@@

# calculate R*
r_star_nom_gbm <- r_star(full_data, method="gbm", uncertainty = T)
r_star_nom_rf <- r_star(full_data, method="rf", uncertainty = T)

# calculate rhat as in posterior package
mon_nom <- monitor_extra(full_data)
rhat <- mon_nom$zfsRhat
rhat1 <- mon_nom$zsRhat
rhat_nom <- tibble(rhat, rhat1) %>% 
  apply(1, max)
```

@@

# calculate R*
r_star_alt_gbm <- r_star(full_data, method="gbm", uncertainty = T)
r_star_alt_rf <- r_star(full_data, method="rf", uncertainty = T)

# calculate rhat as in posterior package
mon_alt <- monitor_extra(full_data)
rhat <- mon_alt$zfsRhat
rhat1 <- mon_alt$zsRhat
rhat_alt <- tibble(rhat, rhat1) %>% 
  apply(1, max)

### 2020-11-09 src/s_discrete.Rmd
cleaning up repo and rerunning rhat experiments

Generate replicates and calculate R* on these
```{r}
f_replicate <- function(L, P0, P1, state_0){
  temp <- f_generate_four(L, P0, P1, state_0)
  a_array <- array(dim=c(L, 4, 1))
  a_array[,,1] <- temp
  a_accuracy <- r_star(a_array)
  mon <- monitor_extra(split_data(a_array))
  return(list(r_star=a_accuracy, r_hat=mon$zfsRhat))
}

@@

saveRDS(tibble(r_star=r_star_vals, r_hat=r_hat_vals), "../data/discrete_r_star_hat.rds")
```

@@

saveRDS(list(P0=P0, P1=P1), "../data/discrete_higherd_matrices.rds")
temp <- readRDS("../data/discrete_higherd_matrices.rds")
P0 <- temp$P0
P1 <- temp$P1
# show first few stationary frequencies for the chains
g <- f_generate_four(100000, P0, P1, 1) %>% 
  as.data.frame() %>% 
  rename(chain_1=V1, chain_2=V2, chain_3=V3, chain_4=V4) %>% 
  mutate(iter=seq_along(chain_1)) %>% 
  melt(id.vars="iter") %>% 
  group_by(variable) %>% 
  summarise("1"=mean(value==1),
            "2"=mean(value==2),
            "3"=mean(value==3),
            "4"=mean(value==4)) %>% 
  pivot_longer(cols="1":"4") %>% 
  mutate(variable=if_else(variable=="chain_1", "1", if_else(variable=="chain_2", "2", if_else(variable=="chain_3", "3", "4")))) %>% 
  ggplot(aes(x=as.factor(name), fill=as.factor(variable), y=value)) +
  geom_col(position = position_dodge()) +
  xlab("State") +
  ylab("Frequency") +
  scale_fill_brewer("Chain", palette = "Dark2") +
  theme(axis.text = element_text(colour="black", size=16),
        axis.title = element_text(colour="black", size=16),
        strip.text.y = element_text(colour="black", size=16),
        title = element_text(colour="black", size=16),
        legend.text = element_text(colour="black", size=16))
ggsave("../output/discrete_stationary_higherd.pdf", g, width = 10, height = 6)
```

@@

saveRDS(tibble(r_star=r_star_vals, r_hat=r_hat_vals), "../data/discrete_r_star_hat_higherd.rds")

>>>>

Generate replicates and calculate R* on these
```{r}
f_replicate <- function(L, P0, P1, state_0){
  temp <- f_generate_four(L, P0, P1, state_0)
  a_array <- array(dim=c(L, 4, 1))
  a_array[,,1] <- temp
  a_accuracy <- r_star(a_array, method = "gbm")
  mon <- monitor_extra(a_array)
  rhat <- max(mon$zsRhat, mon$zfsRhat)
  return(list(r_star=a_accuracy, r_hat=rhat))
}

@@

saveRDS(list(r_star=r_star_vals, r_hat=r_hat_vals), "../data/discrete_r_star_hat.rds")
```

@@

# saveRDS(list(P0=P0, P1=P1), "../data/discrete_higherd_matrices.rds")
temp <- readRDS("../data/discrete_higherd_matrices.rds")
P0 <- temp$P0
P1 <- temp$P1
# show first few stationary frequencies for the chains
g <- f_generate_four(100000, P0, P1, 1) %>% 
  as.data.frame() %>% 
  rename(chain_1=V1, chain_2=V2, chain_3=V3, chain_4=V4) %>% 
  mutate(iter=seq_along(chain_1)) %>% 
  melt(id.vars="iter") %>% 
  group_by(variable) %>% 
  summarise("1"=mean(value==1),
            "2"=mean(value==2),
            "3"=mean(value==3),
            "4"=mean(value==4)) %>% 
  pivot_longer(cols="1":"4") %>% 
  mutate(variable=if_else(variable=="chain_1", "1", if_else(variable=="chain_2", "2", if_else(variable=="chain_3", "3", "4")))) %>% 
  ggplot(aes(x=as.factor(name), fill=as.factor(variable), y=value)) +
  geom_col(position = position_dodge()) +
  xlab("State") +
  ylab("Frequency") +
  scale_fill_brewer("Chain", palette = "Dark2") +
  theme(axis.text = element_text(colour="black", size=16),
        axis.title = element_text(colour="black", size=16),
        strip.text.y = element_text(colour="black", size=16),
        title = element_text(colour="black", size=16),
        legend.text = element_text(colour="black", size=16))
# ggsave("../output/discrete_stationary_higherd.pdf", g, width = 10, height = 6)
```

@@

saveRDS(list(r_star=r_star_vals, r_hat=r_hat_vals), "../data/discrete_r_star_hat_higherd.rds")

### 2020-11-09 src/s_eight_schools.Rmd
cleaning up repo and rerunning rhat experiments

---
title: "R Notebook"
output: html_notebook
---

@@

fit_ncp <- sampling(
  model_ncp, data = eight_schools,
  iter = 2000, chains = 4, seed = 483892929, refresh = 0,
  control = list(adapt_delta = 0.95)
)
```
## Break each chain into two then make flattened data
```{r}
# Centered
x <- rstan::extract(fit_cp, permuted=F)
r_star_centered <- r_star(x, uncertainty = T)
mon_cp <- monitor_extra(split_data(x))

# Noncentered
x <- rstan::extract(fit_ncp, permuted=F)
r_star_noncentered <- r_star(x, uncertainty = T)
mon_ncp <- monitor_extra(split_data(x))

c_df <- tibble(centered=r_star_centered,
               non_centered=r_star_noncentered,
               metric="R*",
               simulations="split")
c1_df <- tibble(centered=mon_cp$zfsRhat[1:10],
                non_centered=mon_ncp$zfsRhat[c(1, 2, 11:18)],
                metric="split-Rhat",
               simulations="split")
```

@@

## Trying with original data (i.e. 4 chains)
```{r}
x <- rstan::extract(fit_cp, permuted=F)
r_star_centered <- r_star(x, split_chains = F, uncertainty = T, method="gbm")
mon_cp <- monitor_extra(fit_cp)
# noncentered
x <- rstan::extract(fit_ncp, permuted=F)
r_star_noncentered <- r_star(x, split_chains = F, uncertainty = T, method="gbm")
mon_ncp <- monitor_extra(fit_ncp)

d_df <- tibble(centered=r_star_centered,
               non_centered=r_star_noncentered,
               metric="R*",
               simulations="unsplit")
d1_df <- tibble(centered=mon_cp$zfsRhat[1:10],
                non_centered=mon_ncp$zfsRhat[c(1, 2, 11:18)],
                metric="split-Rhat",
               simulations="unsplit")
```

Look at Rhat distributions
```{r}
e_df <- c1_df %>% 
  mutate(var_num=seq_along(metric)) %>% 
  bind_rows(d1_df %>% mutate(var_num=seq_along(metric)))
g3 <- e_df %>%
  select(-metric) %>% 
  melt(id.vars=c("simulations", "var_num")) %>% 
  pivot_wider(id_cols=c("variable", "var_num"), values_from = "value", names_from = "simulations") %>% 
  ggplot(aes(x=unsplit, y=split, colour=as.factor(variable))) +
  geom_point() +
  scale_x_continuous(limits=c(NA, 1.035)) +
  scale_y_continuous(limits=c(NA, 1.035)) +
  geom_vline(xintercept = 1.01, linetype=2) +
  geom_hline(yintercept = 1.01, linetype=2) +
  geom_abline(linetype=3) +
  scale_color_grey("Series", labels=c("Centered", "Non-centered")) +
  ylab("Split chains") +
  xlab("Original chains") +
  ggtitle("C.") +
  theme(text = element_text(size=16, colour="black"),
        axis.text = element_text(colour="black", size=10),
        legend.position = c(0.72, 0.3))

@@

# Noncentered
x <- rstan::extract(fit_ncp, permuted=F)
r_star_noncentered <- r_star(x, uncertainty = T, method="rf")
mon_ncp <- monitor_extra(split_data(x))

c_df <- tibble(centered=r_star_centered,
               non_centered=r_star_noncentered,
               metric="R*",
               simulations="split")
c1_df <- tibble(centered=mon_cp$zfsRhat[1:10],
                non_centered=mon_ncp$zfsRhat[c(1, 2, 11:18)],
                metric="split-Rhat",
               simulations="split")
```

```{r}
x <- rstan::extract(fit_cp, permuted=F)
r_star_centered <- r_star(x, split_chains = F, uncertainty = T, method="rf")
mon_cp <- monitor_extra(fit_cp)
# noncentered
x <- rstan::extract(fit_ncp, permuted=F)
r_star_noncentered <- r_star(x, split_chains = F, uncertainty = T, method="rf")
mon_ncp <- monitor_extra(fit_ncp)

d_df <- tibble(centered=r_star_centered,
               non_centered=r_star_noncentered,
               metric="R*",
               simulations="unsplit")
d1_df <- tibble(centered=mon_cp$zfsRhat[1:10],
                non_centered=mon_ncp$zfsRhat[c(1, 2, 11:18)],
                metric="split-Rhat",
               simulations="unsplit")
```

>>>>

---
title: "8 schools"
output: html_notebook
---

@@

fit_ncp <- sampling(
  model_ncp, data = eight_schools,
  iter = 2000, chains = 4, seed = 483892929, refresh = 0,
  control = list(adapt_delta = 0.95)
)
```
## Break each chain into two then make flattened data
```{r}
# Centered
x <- rstan::extract(fit_cp, permuted=F)
r_star_centered <- r_star(x, uncertainty = T, method = "gbm")
mon_cp <- monitor_extra(x)

# Noncentered
x <- rstan::extract(fit_ncp, permuted=F)
r_star_noncentered <- r_star(x, uncertainty = T, method = "gbm")
mon_ncp <- monitor_extra(x)

c_df <- tibble(centered=r_star_centered,
               non_centered=r_star_noncentered,
               metric="R*",
               simulations="split")

# calculate Rhat as max of zfs and zs Rhat
f_rhat_max <- function(mon, ids) {
  rhat <- mon$zfsRhat[ids]
  rhat1 <- mon$zsRhat[ids]
  rhat <- tibble(rhat, rhat1) %>% 
    apply(1, max)
  return(rhat)
}

c1_df <- tibble(centered=f_rhat_max(mon_cp, 1:10),
                non_centered=f_rhat_max(mon_ncp, c(1, 2, 11:18)),
                metric="split-Rhat",
               simulations="split")
```

@@

## Trying with original data (i.e. 4 chains)
```{r}
x <- rstan::extract(fit_cp, permuted=F)
r_star_centered <- r_star(x, split_chains = F, uncertainty = T, method="gbm")

# do calculations using unsplit chains
rhat_cp <- vector(length = 10)
for(i in 1:10) {
  sims_i <- x[, , i]
  zsims_split <- z_scale(sims_i)
  zsplit_rhat <- rhat_rfun(zsims_split)
  sims_centered <- sims_i - median(sims_i)
  sims_folded <- abs(sims_centered)
  zsims_folded_split <- z_scale(sims_folded)
  zfsplit_rhat <- rhat_rfun(zsims_folded_split)
  rhat_cp[i] <- max(zsplit_rhat, zfsplit_rhat)
}

# noncentered
x <- rstan::extract(fit_ncp, permuted=F)
r_star_noncentered <- r_star(x, split_chains = F, uncertainty = T, method="gbm")

ids <- c(1, 2, 11:18)
rhat_ncp <- vector(length = 10)
k <- 1
for(i in ids) {
  sims_i <- x[, , i]
  zsims_split <- z_scale(sims_i)
  zsplit_rhat <- rhat_rfun(zsims_split)
  sims_centered <- sims_i - median(sims_i)
  sims_folded <- abs(sims_centered)
  zsims_folded_split <- z_scale(sims_folded)
  zfsplit_rhat <- rhat_rfun(zsims_folded_split)
  rhat_ncp[k] <- max(zsplit_rhat, zfsplit_rhat)
  k <- k + 1
}

d_df <- tibble(centered=r_star_centered,
               non_centered=r_star_noncentered,
               metric="R*",
               simulations="unsplit")
d1_df <- tibble(centered=rhat_cp,
                non_centered=rhat_ncp,
                metric="split-Rhat",
               simulations="unsplit")
```

Plot two R* distributions
```{r}
g1 <- 
  c_df %>% 
  melt(id.vars=c("metric", "simulations")) %>% 
  ggplot(aes(x=value, fill=as.factor(variable))) +
  geom_histogram(position="identity", alpha=0.8) +
  xlim(0.78, 1.7) +
  scale_fill_grey("Series", labels=c("centered", "non-centered")) +
  xlab(TeX("$R*$")) +
  ylab("Count") +
  geom_vline(xintercept = 1, linetype=2) +
  ggtitle("A.") +
  theme(text = element_text(size=16, colour="black"),
        axis.text = element_text(colour="black"),
        legend.position = "none")
g2 <- 
  d_df %>% 
  melt(id.vars=c("metric", "simulations")) %>% 
  ggplot(aes(x=value, fill=as.factor(variable))) +
  geom_histogram(position="identity", alpha=0.8) +
  scale_fill_grey("Series", labels=c("centered", "non-centered")) +
  xlim(0.78, 1.7) +
  xlab(TeX("$R*$")) +
  ylab("Count") +
  geom_vline(xintercept = 1, linetype=2) +
  ggtitle("B.") +
  theme(text = element_text(size=16, colour="black"),
        axis.text = element_text(colour="black"),
        legend.position = "none")
# a_df <- readRDS("../output/eight_schools_replicates_4_index.rds")
```

Look at Rhat distributions
```{r}
e_df <- c1_df %>% 
  mutate(var_num=seq_along(metric)) %>% 
  bind_rows(d1_df %>% mutate(var_num=seq_along(metric)))
g3 <- e_df %>%
  select(-metric) %>% 
  melt(id.vars=c("simulations", "var_num")) %>% 
  pivot_wider(id_cols=c("variable", "var_num"), values_from = "value", names_from = "simulations") %>% 
  ggplot(aes(x=unsplit, y=split, colour=as.factor(variable))) +
  geom_point() +
  scale_x_continuous(limits=c(NA, 1.02)) +
  scale_y_continuous(limits=c(NA, 1.02)) +
  geom_vline(xintercept = 1.01, linetype=2) +
  geom_hline(yintercept = 1.01, linetype=2) +
  geom_abline(linetype=3) +
  scale_color_grey("Series", labels=c("Centered", "Non-centered")) +
  ylab("Split chains") +
  xlab("Original chains") +
  ggtitle("C.") +
  theme(text = element_text(size=16, colour="black"),
        axis.text = element_text(colour="black", size=10),
        legend.position = c(0.72, 0.3))

@@

# Noncentered
x <- rstan::extract(fit_ncp, permuted=F)
r_star_noncentered <- r_star(x, uncertainty = T, method="rf")

c_df <- tibble(centered=r_star_centered,
               non_centered=r_star_noncentered,
               metric="R*",
               simulations="split")
```

```{r}
x <- rstan::extract(fit_cp, permuted=F)
r_star_centered <- r_star(x, split_chains = F, uncertainty = T, method="rf")
mon_cp <- monitor_extra(fit_cp)
# noncentered
x <- rstan::extract(fit_ncp, permuted=F)
r_star_noncentered <- r_star(x, split_chains = F, uncertainty = T, method="rf")

d_df <- tibble(centered=r_star_centered,
               non_centered=r_star_noncentered,
               metric="R*",
               simulations="unsplit")
```

### 2020-11-09 src/s_gbm_vs_rf.Rmd
cleaning up repo and rerunning rhat experiments

Generate a large correlation matrix
```{r}
rlkj <- function(n, nu, d){
  r_list <- vector(length = n, mode = 'list')
  for(i in 1:n){
    if(d==1)
      r = as.array(1)
    else if(d==2){
      rho <- 2 * rbeta(1, nu, nu) - 1
      r <- matrix(c(1, rho, rho, 1), ncol = 2)
    }else{
      beta <- nu + (d - 2) / 2
      u <- rbeta(1, beta, beta)
      r_12 <- 2 * u - 1
      r <- matrix(c(1, r_12, r_12, 1), ncol = 2)
      for(m in 2:(d - 1)){
        beta <- beta - 0.5
        y <- rbeta(1, m / 2, beta)
        a <- rnorm(m)
        anorm <- sqrt(sum(a^2))
        u <- a / anorm
        w <- sqrt(y) * u
        A <- chol(r)
        z <- w%*%A
        r <-cbind(r, t(z))
        r <- rbind(r, c(z, 1))
      }
    }
    r_list[[i]] <- r
  }
  return(r_list)
}
# d <- 32
# A0 <- rlkj(1, 1, d)[[1]]
# A1 <- rlkj(1, 1, d)[[1]]
# saveRDS(list(A0=A0, A1=A1), "../data/xgbTree_vs_rf_normals.rds")
temp <- readRDS("../data/xgbTree_vs_rf_normals.rds")
A0 <- temp$A0
A1 <- temp$A1

@@

nreplicates <- 20
ds <- c(1, 2, 4, 8, 16, 32)
for(i in seq_along(ds)){
  print(i)
  temp_df <- f_compare_iter(nreplicates, ds[i], A0, A1)
  if(i==1)
    big_df <- temp_df
  else
    big_df <- big_df %>% bind_rows(temp_df)
}
saveRDS(big_df, "../data/gbm_vs_rf_normals_results.rds")

@@

r_star_opts <- map_dbl(ds[2:length(ds)], ~f_optimal_normals(10000, A0, A1, .))
r_star_opts <- c(1, r_star_opts)
r_star_opts <- tibble(d=ds, optimal=r_star_opts)
temp_df <- 
  big_df %>%
  left_join(r_star_opts) %>% 
  select(-contains("time")) %>% 
  melt(id.vars=c("iter", "d")) %>% 
  group_by(variable, d) %>% 
  summarise(lower=quantile(value, 0.25),
            middle=quantile(value, 0.5),
            upper=quantile(value, 0.75)) %>% 
  ungroup() %>% 
  mutate(variable=as.character(variable)) %>% 
  mutate(variable=if_else(variable=="rstar_rf", "rf", if_else(variable=="rstar_gbm", "gbm", "optimal")))
saveRDS(temp_df, "../data/gbm_vs_rf_normals_results_optimals.rds")
temp_df <- readRDS("../data/gbm_vs_rf_normals_results_optimals.rds")
g <- ggplot(data=filter(temp_df, variable!="optimal"),
       aes(x=as.factor(d), y=middle, colour=as.factor(variable))) +
  geom_pointrange(aes(ymin=lower, ymax=upper),
                  position = position_dodge2(width=0.3)) +
  geom_hline(yintercept = 1, linetype=2) +
  geom_point(data=filter(temp_df, variable=="optimal")) +
  scale_color_brewer("Method", palette = "Dark2") +
  xlab("# dimensions") +
  ylab("R*") +
  theme(text = element_text(size=16, colour="black"),
        axis.text = element_text(colour="black"))
ggsave("../output/gbm_vs_rf_normal.pdf", g, width = 10, height = 6)
```

Same but increase fatness of tails
```{r}
library(LaplacesDemon)
f_data_student_ts <- function(nu1, nu2, d, A0, niter=2000){
  x <- array(dim=c(niter, 4, d))
  # ensures variance stays same
  sigma1 <- A0[1:d, 1:d] * (nu1 - 2) / nu1
  sigma2 <- A0[1:d, 1:d] * (nu2 - 2) / nu2
  if(d > 1){
    for(i in 1:3)
      x[, i, ] <- mvtnorm::rmvt(niter, sigma=sigma1, df=nu1, delta=rep(0, d))
    x[ , 4, ] <- mvtnorm::rmvt(niter, sigma=sigma2, df=nu2, delta=rep(0, d))
  }else{
    for(i in 1:3)
      x[, i, ] <- LaplacesDemon::rst(niter, 0, sqrt(sigma1), nu1)
    x[ , 4, ] <- LaplacesDemon::rst(niter, 0, sqrt(sigma2), nu2)
  }
  return(x)
}

@@

nreplicates <- 20
nu1 <- 3
nu2 <- c(4, 8, 16, 32)
ds <- c(1, 2, 4, 8, 16, 32)
m_parameters <- expand_grid(nu2, ds)

@@

f_optimal_studentts <- function(niter, nu1, nu2, d, A0){
  x <- f_data_student_ts(nu1, nu2, d, A0, niter = niter)
  sigma1 <- A0[1:d, 1:d] * (nu1 - 2) / nu1
  sigma2 <- A0[1:d, 1:d] * (nu2 - 2) / nu2
  if(d > 1){
    f_first <- function(x) dmvt(x, rep(0, d), sigma1, nu1)
    f_second <- function(x) dmvt(x, rep(0, d), sigma2, nu2)
  }else{
    f_first <- function(x) dst(x, 0, sigma1, nu1)
    f_second <- function(x) dst(x, 0, sigma2, nu2)
  }
  m_class <- matrix(nrow = dim(x)[1], ncol = dim(x)[2])
  for(i in 1:dim(x)[1]){
    for(j in 1:dim(x)[2]){
      prob0 <- f_first(x[i, j,])
      prob1 <- f_second(x[i, j,])
      if(prob0 > prob1)
        m_class[i, j] <- 0
      else
        m_class[i, j] <- 1
    }
  }
  a0 <- 1/3 * 1/3 * (mean(m_class[, 1]==0) + mean(m_class[, 2]==0) + mean(m_class[, 3]==0))
  a1 <- mean(m_class[, 4]==1)
  return(1/4 * (3 * a0 + a1) / 0.25)
}

r_star_opts1 <- map_dbl(seq(1, nrow(m_parameters), 1), ~f_optimal_studentts(1000, nu1, m_parameters$nu2[.], m_parameters$ds[.], A0))

r_star_opts <- tibble(d=m_parameters$ds, nu2=m_parameters$nu2, optimal=r_star_opts1)
temp_df <- 
  big_df %>%
  left_join(r_star_opts) %>% 
  select(-contains("time")) %>% 
  select(-one_of("nu1")) %>% 
  melt(id.vars=c("iter", "d", "nu2")) %>% 
  group_by(variable, d, nu2) %>% 
  summarise(lower=quantile(value, 0.25),
            middle=quantile(value, 0.5),
            upper=quantile(value, 0.75)) %>% 
  ungroup() %>% 
  mutate(variable=as.character(variable)) %>% 
  mutate(variable=if_else(variable=="rstar_rf", "rf", if_else(variable=="rstar_gbm", "gbm", "optimal")))
saveRDS(temp_df, "../data/gbm_vs_rf_studentts_results_optimal.rds")
temp_df <- readRDS("../data/gbm_vs_rf_studentts_results_optimal.rds")
g2 <- ggplot(data=filter(temp_df, variable!="optimal"),
       aes(x=as.factor(nu2), y=middle, colour=as.factor(variable))) +
  geom_pointrange(aes(ymin=lower, ymax=upper),
                  position = position_dodge2(width=0.3)) +
  geom_hline(yintercept = 1, linetype=2) +
  geom_point(data=filter(temp_df, variable=="optimal")) +
  scale_color_brewer("Method", palette = "Dark2") +
  xlab("# degrees of freedom") +
  ylab("R*") +
  theme(text = element_text(size=16, colour="black"),
        axis.text = element_text(colour="black")) +
  facet_wrap(~d)
ggsave("../output/gbm_vs_rf_studentt.pdf", g2, width = 12, height = 8)
```

>>>>

# Different covariance matrices
```{r}
rlkj <- function(n, nu, d){
  r_list <- vector(length = n, mode = 'list')
  for(i in 1:n){
    if(d==1)
      r = as.array(1)
    else if(d==2){
      rho <- 2 * rbeta(1, nu, nu) - 1
      r <- matrix(c(1, rho, rho, 1), ncol = 2)
    }else{
      beta <- nu + (d - 2) / 2
      u <- rbeta(1, beta, beta)
      r_12 <- 2 * u - 1
      r <- matrix(c(1, r_12, r_12, 1), ncol = 2)
      for(m in 2:(d - 1)){
        beta <- beta - 0.5
        y <- rbeta(1, m / 2, beta)
        a <- rnorm(m)
        anorm <- sqrt(sum(a^2))
        u <- a / anorm
        w <- sqrt(y) * u
        A <- chol(r)
        z <- w%*%A
        r <-cbind(r, t(z))
        r <- rbind(r, c(z, 1))
      }
    }
    r_list[[i]] <- r
  }
  return(r_list)
}
# d <- 32
# A0 <- rlkj(1, 1, d)[[1]]
# A1 <- rlkj(1, 1, d)[[1]]
# saveRDS(list(A0=A0, A1=A1), "../data/xgbTree_vs_rf_normals.rds")
temp <- readRDS("../data/xgbTree_vs_rf_normals.rds")
A0 <- temp$A0
A1 <- temp$A1

@@

# Run replicates with one chain with a different cov matrix
nreplicates <- 20
ds <- c(1, 2, 4, 8, 16, 32)
for(i in seq_along(ds)){
  print(i)
  temp_df <- f_compare_iter(nreplicates, ds[i], A0, A1)
  if(i==1)
    big_df <- temp_df
  else
    big_df <- big_df %>% bind_rows(temp_df)
}
saveRDS(big_df, "../data/gbm_vs_rf_normals_results.rds")

@@

r_star_opts <- map_dbl(ds[2:length(ds)], ~f_optimal_normals(10000, A0, A1, .))
r_star_opts <- c(1, r_star_opts)
r_star_opts <- tibble(d=ds, optimal=r_star_opts)

# plot results
temp_df <- 
  big_df %>%
  left_join(r_star_opts) %>% 
  select(-contains("time")) %>% 
  melt(id.vars=c("iter", "d")) %>% 
  group_by(variable, d) %>% 
  summarise(lower=quantile(value, 0.25),
            middle=quantile(value, 0.5),
            upper=quantile(value, 0.75)) %>% 
  ungroup() %>% 
  mutate(variable=as.character(variable)) %>% 
  mutate(variable=if_else(variable=="rstar_rf", "rf", if_else(variable=="rstar_gbm", "gbm", "optimal")))
saveRDS(temp_df, "../data/gbm_vs_rf_normals_results_optimals.rds")
temp_df <- readRDS("../data/gbm_vs_rf_normals_results_optimals.rds") %>% 
  mutate(variable=as.character(variable)) %>% 
  mutate(variable=if_else(variable=="gbm", "GBM",
                          if_else(variable=="rf", "RF", "Optimal"))) %>% 
  mutate(variable=fct_relevel(as.factor(variable), "GBM", "RF", "Optimal"))
g <- ggplot(temp_df,
       aes(x=as.factor(d), y=middle, colour=as.factor(variable),
           shape=as.factor(variable))) +
  geom_pointrange(aes(ymin=lower, ymax=upper),
                  position = position_dodge2(width=0.3)) +
  geom_hline(yintercept = 1, linetype=2) +
  scale_color_grey("Method", start=0.8, end=0) +
  scale_shape("Method") +
  xlab("# dimensions") +
  ylab("R*") +
  theme(text = element_text(size=16, colour="black"),
        axis.text = element_text(colour="black"))
ggsave("../output/gbm_vs_rf_normal.pdf", g, width = 10, height = 6)
```

# Tail fatness
```{r}
library(LaplacesDemon)
f_data_student_ts <- function(nu1, nu2, d, A0, niter=2000){
  x <- array(dim=c(niter, 4, d))
  # ensures variance stays same
  sigma1 <- A0[1:d, 1:d] * (nu1 - 2) / nu1
  sigma2 <- A0[1:d, 1:d] * (nu2 - 2) / nu2
  if(d > 1){
    for(i in 1:3)
      x[, i, ] <- mvtnorm::rmvt(niter, sigma=sigma1, df=nu1, delta=rep(0, d))
    x[ , 4, ] <- mvtnorm::rmvt(niter, sigma=sigma2, df=nu2, delta=rep(0, d))
  }else{
    for(i in 1:3)
      x[, i, ] <- LaplacesDemon::rst(niter, 0, sqrt(sigma1), nu1)
    x[ , 4, ] <- LaplacesDemon::rst(niter, 0, sqrt(sigma2), nu2)
  }
  return(x)
}

@@

# Run replicates
nreplicates <- 20
nu1 <- 3
nu2 <- c(4, 8, 16, 32)
ds <- c(1, 2, 4, 8, 16, 32)
m_parameters <- expand_grid(nu2, ds)

@@

# calculate optimal classification rates
f_optimal_studentts <- function(niter, nu1, nu2, d, A0){
  x <- f_data_student_ts(nu1, nu2, d, A0, niter = niter)
  sigma1 <- A0[1:d, 1:d] * (nu1 - 2) / nu1
  sigma2 <- A0[1:d, 1:d] * (nu2 - 2) / nu2
  if(d > 1){
    f_first <- function(x) dmvt(x, rep(0, d), sigma1, nu1)
    f_second <- function(x) dmvt(x, rep(0, d), sigma2, nu2)
  }else{
    f_first <- function(x) dst(x, 0, sigma1, nu1)
    f_second <- function(x) dst(x, 0, sigma2, nu2)
  }
  m_class <- matrix(nrow = dim(x)[1], ncol = dim(x)[2])
  for(i in 1:dim(x)[1]){
    for(j in 1:dim(x)[2]){
      prob0 <- f_first(x[i, j,])
      prob1 <- f_second(x[i, j,])
      if(prob0 > prob1)
        m_class[i, j] <- 0
      else
        m_class[i, j] <- 1
    }
  }
  a0 <- 1/3 * 1/3 * (mean(m_class[, 1]==0) + mean(m_class[, 2]==0) + mean(m_class[, 3]==0))
  a1 <- mean(m_class[, 4]==1)
  return(1/4 * (3 * a0 + a1) / 0.25)
}

r_star_opts1 <- map_dbl(seq(1, nrow(m_parameters), 1), ~f_optimal_studentts(10000, nu1, m_parameters$nu2[.], m_parameters$ds[.], A0))

# plot results
r_star_opts <- tibble(d=m_parameters$ds, nu2=m_parameters$nu2, optimal=r_star_opts1)
big_df <- readRDS("../data/gbm_vs_rf_studentts_results.rds")
temp_df <- 
  big_df %>%
  left_join(r_star_opts) %>% 
  select(-contains("time")) %>% 
  select(-one_of("nu1")) %>% 
  melt(id.vars=c("iter", "d", "nu2")) %>% 
  group_by(variable, d, nu2) %>% 
  summarise(lower=quantile(value, 0.25),
            middle=quantile(value, 0.5),
            upper=quantile(value, 0.75)) %>% 
  ungroup() %>% 
  mutate(variable=as.character(variable)) %>% 
  mutate(variable=if_else(variable=="rstar_rf", "rf", if_else(variable=="rstar_gbm", "gbm", "optimal")))
saveRDS(temp_df, "../data/gbm_vs_rf_studentts_results_optimal.rds")
temp_df <- readRDS("../data/gbm_vs_rf_studentts_results_optimal.rds") %>% 
  mutate(variable=as.character(variable)) %>% 
  mutate(variable=if_else(variable=="gbm", "GBM",
                          if_else(variable=="rf", "RF", "Optimal"))) %>% 
  mutate(variable=fct_relevel(as.factor(variable), "GBM", "RF", "Optimal"))
g2 <- ggplot(data=temp_df,
       aes(x=as.factor(nu2), y=middle, colour=as.factor(variable),
           shape=as.factor(variable))) +
  geom_pointrange(aes(ymin=lower, ymax=upper),
                  position = position_dodge2(width=0.3)) +
  geom_hline(yintercept = 1, linetype=2) +
  scale_color_grey("Method", start=0.8, end=0) +
  scale_shape("Method") +
  xlab("# degrees of freedom") +
  ylab("R*") +
  theme(text = element_text(size=16, colour="black"),
        axis.text = element_text(colour="black")) +
  facet_wrap(~d)
ggsave("../output/gbm_vs_rf_studentt.pdf", g2, width = 12, height = 8)
```
